Titeln på avhandlingens språk

### 2022-12-19 publications/2022Ainur/introduction.tex
glossary fixes

Over the last years, several small- to mid-scale testbeds have emerged that leverage a large degree of freedom with respect to hardware and software, for example, the 
\begin{enumerate*}[itemjoin={{, }}, itemjoin*={{, and }}]
    \item \acs*{COSMOS}
    \item \acs*{POWDER}
    \item \acs*{ARA}
    \item Drexel Grid \gls{SDR}
\end{enumerate*} testbeds.
\acs{COSMOS}~(\emph{\acl{COSMOS}})\acused{COSMOS} is a testbed spanning an area of roughly \num{1} square mile (\SI{2.6}{\kilo\meter\squared}) featuring \acp{SDR}, \si{\milli\meter}-wave equipment, optical fibers, cloud integration, and compute for core network functionality and application data processing~\cite{Cosmos1,cosmos2}.
It contains over \num{200} rooftop, intermediate, and mobile nodes, and is controlled and managed by a central node.
\gls{COSMOS} relies on the \gls{OMF} (originally developed for ORBIT~\cite{orbit}), and employs the \gls{OEDL}, a domain-specific imperative language based on Ruby, for experiment development and definition.

\acs{POWDER}~(\emph{\acl{POWDER}})\acused{POWDER} promises research on wireless and mobile networks with a level of programmability down to the waveform~\cite{powder}.
The testbed spans a \SI{15}{\kilo\meter\squared} area and features about \num{15} fixed programmable radio nodes, based on off-the-shelves \acp{SDR} and featuring edge-like compute capabilities and integration with cloud resources, which interact with \num{50} mobile nodes. 
\gls{POWDER} experiments are defined and developed in \emph{profiles}, which correspond to \gls{VM} images containing the necessary software and configurations.
These profiles are defined through using \gls{RSpec}\footnote{\url{https://groups.geni.net/geni/wiki/GENIExperimenter/RSpecs}} documents.

The \gls{ARA} platform is an at-scale testbed for wireless research spanning a rural area with a diameter of over \SI{60}{\kilo\meter} in Iowa~\cite{zhang2022ara}.
Its core goal is the study and deployment of advanced wireless platforms and technologies in real-world agricultural and rural settings.
It includes a broad range of wireless platforms ranging from low-\gls{UHF} massive \gls{MIMO} to \unit{\milli\meter}Wave, deployed through both \acp{SDR} and programmable \gls{COTS} radios, as well as automated ground vehicles, cameras and sensors.
\gls{ARA}'s software stack, ARASoft, is based upon the highly flexible and powerful \gls{CHI} software suite from the Chameleon testbed project~\cite{keahey2020lessons}, which in turn is based on the widely adopted OpenStack~\cite{openstack} cloud-computing framework.
This affords the \gls{ARA} testbed a high degree of flexibility, as well as lowers the learning curve for new contributors and users.

Finally, the \emph{Drexel Grid \gls{SDR} Testbed} features \acp{SDR} that connect either over-the-air, through a channel emulator, or over a combination of the two, to facilitate realistic and reproducible experimentation~\cite{DrexelGrid}.
Primarily intended for \gls{SDR}-centric research, it does not integrate any core, cloud or edge components.
However, the testbed extensively employs the \gls{LXC} runtime for the deployment of both experimental code and \gls{SDR} software, which affords users great freedom when it comes to the development of experiments.

Experimentation is key to to fully understanding the implications of next-generation wireless systems, cloud, and edge computing paradigms, and thus more of these testbeds are sure to emerge in the near future.
Yet, little work has so far been devoted to general-purpose, hardware-agnostic software frameworks for the management and automation of such systems.
Existing platforms implement their own, ad-hoc software solutions which are not compatible with other testbeds, and in many cases are not even compatible with reigning cloud-native standards.
This is, for instance, the case with \gls{COSMOS} and \gls{POWDER}; their reliance on domain-specific languages limits their integration with cloud-native solutions, which generally build upon general-purpose languages such as Python and Go.
These testbeds further leverage virtualization technology based on \acp{VM} instead of more lightweight and edge-compatible solutions such as containers.

To the best of our knowledge, CloudRAFT~\cite{cloudraft} is the only work to tackle (to a certain extent) this challenge.
CloudRAFT corresponds to a cloud-based framework for mobile network experimentation, with a focus on simplifying the management of testbed resources.
The goal of this project is to integrate, coordinate, share, and improve upon existing testbeds; and employs pre-built \acp{VM} containing the necessary software for experiments.
Although it provides some automation for testbed resource provisioning and experiment execution, its focus is largely rather on the sharing and partitioning of testbed systems.
Testbeds currently working with CloudRAFT include a variety of domain-specific setups, including \iac{SDR}-based testbed as a well as a ground vehicular robot for mobility-related experimentation.

@@

% \todo[inline]{
%     COSMOS: Experimenters work on these resources using the \ac*{OMF}~\cite{omf} tool, originally leveraged for ORBIT~\cite{orbit}, and \gls{SSH} sessions, thus allowing for free-form scripting of experimental procedures.

@@

%     POWDER: To abstract and redistribute resources and \acp{API}, \gls{POWDER} leverages the concept of \emph{profiles}, corresponding shareable packages containing experimental configurations and descriptions\cite{emulab}.
%     Automation and web interface control are limited to some test experiments, and thus regular users must interact with the platform through \gls{SSH}.

>>>>

Over the last years, several small- to mid-scale testbeds have emerged that leverage a large degree of freedom with respect to hardware and software, for example, the 
\begin{enumerate*}[itemjoin={{, }}, itemjoin*={{, and }}]
    \item \gls{COSMOS}
    \item \gls{POWDER}
    \item \gls{ARA}
    \item Drexel Grid \gls{SDR}
\end{enumerate*} testbeds.
\gls{COSMOS}~(\emph{\gls{COSMOS}}) is a testbed spanning an area of roughly \num{1} square mile (\SI{2.6}{\kilo\meter\squared}) featuring \glspl{SDR}, \si{\milli\meter}-wave equipment, optical fibers, cloud integration, and compute for core network functionality and application data processing~\cite{Cosmos1,cosmos2}.
It contains over \num{200} rooftop, intermediate, and mobile nodes, and is controlled and managed by a central node.
\gls{COSMOS} relies on the \gls{OMF} (originally developed for ORBIT~\cite{orbit}), and employs the \gls{OEDL}, a domain-specific imperative language based on Ruby, for experiment development and definition.

\gls{POWDER}~(\emph{\gls{POWDER}}) promises research on wireless and mobile networks with a level of programmability down to the waveform~\cite{powder}.
The testbed spans a \SI{15}{\kilo\meter\squared} area and features about \num{15} fixed programmable radio nodes, based on off-the-shelves \glspl{SDR} and featuring edge-like compute capabilities and integration with cloud resources, which interact with \num{50} mobile nodes.
\gls{POWDER} experiments are defined and developed in \emph{profiles}, which correspond to \gls{VM} images containing the necessary software and configurations.
These profiles are defined through using \gls{RSpec}\footnote{\url{https://groups.geni.net/geni/wiki/GENIExperimenter/RSpecs}} documents.

The \gls{ARA} platform is an at-scale testbed for wireless research spanning a rural area with a diameter of over \SI{60}{\kilo\meter} in Iowa~\cite{zhang2022ara}.
Its core goal is the study and deployment of advanced wireless platforms and technologies in real-world agricultural and rural settings.
It includes a broad range of wireless platforms ranging from low-\gls{UHF} massive \gls{MIMO} to \unit{\milli\meter}Wave, deployed through both \glspl{SDR} and programmable \gls{COTS} radios, as well as automated ground vehicles, cameras and sensors.
\gls{ARA}'s software stack, ARASoft, is based upon the highly flexible and powerful \gls{CHI} software suite from the Chameleon testbed project~\cite{keahey2020lessons}, which in turn is based on the widely adopted OpenStack~\cite{openstack} cloud-computing framework.
This affords the \gls{ARA} testbed a high degree of flexibility, as well as lowers the learning curve for new contributors and users.

Finally, the \emph{Drexel Grid \gls{SDR} Testbed} features \glspl{SDR} that connect either over-the-air, through a channel emulator, or over a combination of the two, to facilitate realistic and reproducible experimentation~\cite{DrexelGrid}.
Primarily intended for \gls{SDR}-centric research, it does not integrate any core, cloud or edge components.
However, the testbed extensively employs the \gls{LXC} runtime for the deployment of both experimental code and \gls{SDR} software, which affords users great freedom when it comes to the development of experiments.

Experimentation is key to to fully understanding the implications of next-generation wireless systems, cloud, and edge computing paradigms, and thus more of these testbeds are sure to emerge in the near future.
Yet, little work has so far been devoted to general-purpose, hardware-agnostic software frameworks for the management and automation of such systems.
Existing platforms implement their own, ad-hoc software solutions which are not compatible with other testbeds, and in many cases are not even compatible with reigning cloud-native standards.
This is, for instance, the case with \gls{COSMOS} and \gls{POWDER}; their reliance on domain-specific languages limits their integration with cloud-native solutions, which generally build upon general-purpose languages such as Python and Go.
These testbeds further leverage virtualization technology based on \glspl{VM} instead of more lightweight and edge-compatible solutions such as containers.

To the best of our knowledge, CloudRAFT~\cite{cloudraft} is the only work to tackle (to a certain extent) this challenge.
CloudRAFT corresponds to a cloud-based framework for mobile network experimentation, with a focus on simplifying the management of testbed resources.
The goal of this project is to integrate, coordinate, share, and improve upon existing testbeds; and employs pre-built \glspl{VM} containing the necessary software for experiments.
Although it provides some automation for testbed resource provisioning and experiment execution, its focus is largely rather on the sharing and partitioning of testbed systems.
Testbeds currently working with CloudRAFT include a variety of domain-specific setups, including \pgls{SDR}-based testbed as a well as a ground vehicular robot for mobility-related experimentation.

@@

% \todo[inline]{
%     COSMOS: Experimenters work on these resources using the \gls{OMF}~\cite{omf} tool, originally leveraged for ORBIT~\cite{orbit}, and \gls{SSH} sessions, thus allowing for free-form scripting of experimental procedures.

@@

%     POWDER: To abstract and redistribute resources and \glspl{API}, \gls{POWDER} leverages the concept of \emph{profiles}, corresponding shareable packages containing experimental configurations and descriptions\cite{emulab}.
%     Automation and web interface control are limited to some test experiments, and thus regular users must interact with the platform through \gls{SSH}.

### 2022-12-19 publications/2022Ainur/software_layers/cloud.tex
glossary fixes

The spawned \gls{EC2} instances are created from a preconfigured \gls{AMI}, the ID of which must be passed to the \mintinline{python}{CloudInstances} instance at runtime.
This \gls{AMI} can be either be one of the default \acp{AMI} provided by \acp{AWS} --- e.g.\ a plain Ubuntu or CentOS \gls{AMI} --- or a custom \gls{AMI} which has been preconfigured by the user.
Note that although the \mintinline{python}{CloudInstances} instance places no restrictions on the \gls{AMI} used for deploying (with the exception that it needs to be a Linux-based \gls{AMI}), components such as the VPN layer (see \cref{sec:vpn}) depending on these cloud instances might have requirements of their own.

>>>>

The spawned \gls{EC2} instances are created from a preconfigured \gls{AMI}, the ID of which must be passed to the \mintinline{python}{CloudInstances} instance at runtime.
This \gls{AMI} can be either be one of the default \glspl{AMI} provided by \glspl{AWS} --- e.g.\ a plain Ubuntu or CentOS \gls{AMI} --- or a custom \gls{AMI} which has been preconfigured by the user.
Note that although the \mintinline{python}{CloudInstances} instance places no restrictions on the \gls{AMI} used for deploying (with the exception that it needs to be a Linux-based \gls{AMI}), components such as the VPN layer (see \cref{sec:vpn}) depending on these cloud instances might have requirements of their own.

### 2022-12-19 publications/2022Ainur/software_layers/physical.tex
glossary fixes

In order to ensure flexibility, we make no assumptions on the wired connections between the data hosts or radios.
On the other hand, an arbitrary data plane could require wired connections between two hosts, or a host and a radio.
Therefore, we use a managed switch at the heart of the physical layer and connect all the hosts and radios to it.
According to the data network description, the physical layer then creates the required \acp{VLAN} in the switch.

At instantiation time, this class first configures ethernet connections by creating \acp{VLAN} in the workload data network switch.
It then triggers the deployment of potential wireless connections using \mintinline{python}{WirelessConnection} objects.

A set of \mintinline{python}{WirelessConnection} instances must be provided to the constructor for deployment.
This class can be used as a context manager in a \mintinline{python}{with} statement; at the end of the block, it tears down all connections and cleans the added \acp{VLAN} from the switch.

@@

\begin{description}[style=nextline]
    \item[SDR-based WiFi Network]
    The \mintinline{python}{SDRWifi} class handles the configuration and deployment of a software-defined IEEE 802.11 network using the Mangocomm software package and its supported software-defined radios \footnote{See https://mangocomm.com/802.11-mac-phy/}.
    It lunches a Docker container on the same machine to communicate with the \acp{SDR} over the control network.
    This communication is for the configuration and monitoring the WiFi nodes.
    In a WiFi network, a radio is either \emph{station} or \emph{access point}.
    To initialize a \mintinline{python}{SDRWifi} class, the user is required to provide two lists containing the access points and station nodes instances as described below.

@@

    When a WiFi network is up and an station is connected, the data network interfaces of the radios behave as if an ethernet cable is connecting them.
    Thus, the client hosts and the server hosts can be on the same ip layer subnets.
    \begin{figure}[t]
        \centering
        \includegraphics[width=0.8\linewidth]{figures/wifi.png}
        \caption{An example of WiFi network topology in data network and its corresponding \acp{VLAN}}
        \label{fig:wifi}
    \end{figure}

@@

    \begin{figure}[t]
        \centering
        \includegraphics[width=0.8\linewidth]{figures/lte.png}
        \caption{An example of \gls{LTE} data network topology in data network and its corresponding \acp{VLAN}}
        \label{fig:lte}
    \end{figure}

>>>>

In order to ensure flexibility, we make no assumptions on the wired connections between the data hosts or radios.
On the other hand, an arbitrary data plane could require wired connections between two hosts, or a host and a radio.
Therefore, we use a managed switch at the heart of the physical layer and connect all the hosts and radios to it.
According to the data network description, the physical layer then creates the required \glspl{VLAN} in the switch.

At instantiation time, this class first configures ethernet connections by creating \glspl{VLAN} in the workload data network switch.
It then triggers the deployment of potential wireless connections using \mintinline{python}{WirelessConnection} objects.

A set of \mintinline{python}{WirelessConnection} instances must be provided to the constructor for deployment.
This class can be used as a context manager in a \mintinline{python}{with} statement; at the end of the block, it tears down all connections and cleans the added \glspl{VLAN} from the switch.

@@

\begin{description}[style=nextline]
    \item[SDR-based WiFi Network]
    The \mintinline{python}{SDRWifi} class handles the configuration and deployment of a software-defined IEEE 802.11 network using the Mangocomm software package and its supported software-defined radios \footnote{See https://mangocomm.com/802.11-mac-phy/}.
    It lunches a Docker container on the same machine to communicate with the \glspl{SDR} over the control network.
    This communication is for the configuration and monitoring the WiFi nodes.
    In a WiFi network, a radio is either \emph{station} or \emph{access point}.
    To initialize a \mintinline{python}{SDRWifi} class, the user is required to provide two lists containing the access points and station nodes instances as described below.

@@

    When a WiFi network is up and an station is connected, the data network interfaces of the radios behave as if an ethernet cable is connecting them.
    Thus, the client hosts and the server hosts can be on the same ip layer subnets.
    \begin{figure}[t]
        \centering
        \includegraphics[width=0.8\linewidth]{figures/wifi.png}
        \caption{An example of WiFi network topology in data network and its corresponding \glspl{VLAN}}
        \label{fig:wifi}
    \end{figure}

@@

    \begin{figure}[t]
        \centering
        \includegraphics[width=0.8\linewidth]{figures/lte.png}
        \caption{An example of \gls{LTE} data network topology in data network and its corresponding \glspl{VLAN}}
        \label{fig:lte}
    \end{figure}

### 2022-12-19 publications/2022CLEAVE/cleave.tex
glossary fixes

In this section, we detail our \gls{CLEAVE} framework for the performance evaluation of \aclp{NCS}, with an emphasis on edge deployments.
Our design follows a virtualized emulation approach; providing a framework for the real-time \emph{emulation} of physical control systems and the easy implementation of softwarized controllers interacting with \emph{real} networks.
\gls{CLEAVE} is thus a platform with an easy-to-use \gls{API} on which a multitude of \gls{NCS} types can be emulated.
This allows for great flexibility, as the core components of the \gls{NCS} can be switched out while maintaining the realism of the network, the most complex and limiting component of edge \acp{NCS}.

@@

Implementation-wise, \gls{CLEAVE} is built on top of the well-established and stable \emph{Twisted}\footnote{\url{https://twistedmatrix.com}} asynchronous programming and networking framework.
The aforementioned core components (\emph{State}, \emph{Controller}, \emph{Sensors}, and \emph{Actuators}) are provided as \acp{ABC} with core functionality that users must extend for their specific use cases and \acp{NCS}.
The framework makes no assumptions about the inner workings of the user implementations of these classes other than the interfaces they expose.
Users are therefore free to implement arbitrarily complex functionality in these components.
On the other hand, \gls{CLEAVE} handles network communication and metric collection autonomously; users only need to select the desired transport protocol (we do however provide an \gls{API} for extending the available transports).

@@

\subsection{Configuring and Executing \iac{NCS}}

@@

Although the core design of \gls{CLEAVE} follows a single-loop approach, multi-loop scenarios can be easily implemented through the combination of multiple \gls{CLEAVE} instances.
The framework does not make any assumptions about the individual components other than the signatures of methods on the \acp{ABC} that must be extended and implemented.
Combined with the Python emulation configuration scripts, this makes development of more complex scenarios merely a matter of implementing the desired functionality within the framework itself.

>>>>

In this section, we detail our \gls{CLEAVE} framework for the performance evaluation of \glspl{NCS}, with an emphasis on edge deployments.
Our design follows a virtualized emulation approach; providing a framework for the real-time \emph{emulation} of physical control systems and the easy implementation of softwarized controllers interacting with \emph{real} networks.
\gls{CLEAVE} is thus a platform with an easy-to-use \gls{API} on which a multitude of \gls{NCS} types can be emulated.
This allows for great flexibility, as the core components of the \gls{NCS} can be switched out while maintaining the realism of the network, the most complex and limiting component of edge \glspl{NCS}.

@@

Implementation-wise, \gls{CLEAVE} is built on top of the well-established and stable \emph{Twisted}\footnote{\url{https://twistedmatrix.com}} asynchronous programming and networking framework.
The aforementioned core components (\emph{State}, \emph{Controller}, \emph{Sensors}, and \emph{Actuators}) are provided as \glspl{ABC} with core functionality that users must extend for their specific use cases and \glspl{NCS}.
The framework makes no assumptions about the inner workings of the user implementations of these classes other than the interfaces they expose.
Users are therefore free to implement arbitrarily complex functionality in these components.
On the other hand, \gls{CLEAVE} handles network communication and metric collection autonomously; users only need to select the desired transport protocol (we do however provide an \gls{API} for extending the available transports).

@@

\subsection{Configuring and Executing \pgls{NCS}}

@@

Although the core design of \gls{CLEAVE} follows a single-loop approach, multi-loop scenarios can be easily implemented through the combination of multiple \gls{CLEAVE} instances.
The framework does not make any assumptions about the individual components other than the signatures of methods on the \glspl{ABC} that must be extended and implemented.
Combined with the Python emulation configuration scripts, this makes development of more complex scenarios merely a matter of implementing the desired functionality within the framework itself.

### 2022-12-19 publications/2022CLEAVE/conclusion.tex
glossary fixes

We validate the utility of this tool through an example use case approximating a proposed edge deployment of inverted pendula control loops co-located with video analytics services.
We argue such a use case represents a realistic scenario and appropriate benchmark for the tool, since
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item the inverted pendulum plant is ubiquitous in \gls{NCS} research
    \item similar setups exist in real-world industrial use
    \item video analytics has long been proposed as a ``killer app'' for edge computing.
\end{enumerate*}
Our results showcase the ability of the framework to extract relevant metrics relating to the stability of the control system, as well as on the performance of the underlying network link.
We believe \gls{CLEAVE} represents an important step towards enabling inexpensive and low-complexity scalable research for real-world deployment of edge-bound \acp{NCS}.

There is still, however, work to be done.
We are extending the number of plant and controller implementations on the framework, with the goal of creating an open library of \acp{NCS} to share with the community.
At the moment, the interactions of \gls{CLEAVE} and tools such as Docker are still quite superficial.
Our goal is to achieve a much tighter integration, e.g.\ by providing the toolkit as pre-packaged container images.
Finally, the validity of the results obtained by the framework will have to be verified through more thorough, realistic scenarios than what we have been able to show in this work.
In particular, we intend to perform large-scale experimentation targetting 5G cellular deployments, as this technology is set to become the backbone of edge networks in the near future.

>>>>

We validate the utility of this tool through an example use case approximating a proposed edge deployment of inverted pendula control loops co-located with video analytics services.
We argue such a use case represents a realistic scenario and appropriate benchmark for the tool, since
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item the inverted pendulum plant is ubiquitous in \gls{NCS} research
    \item similar setups exist in real-world industrial use
    \item video analytics has long been proposed as a ``killer app'' for edge computing.
\end{enumerate*}
Our results showcase the ability of the framework to extract relevant metrics relating to the stability of the control system, as well as on the performance of the underlying network link.
We believe \gls{CLEAVE} represents an important step towards enabling inexpensive and low-complexity scalable research for real-world deployment of edge-bound \glspl{NCS}.

There is still, however, work to be done.
We are extending the number of plant and controller implementations on the framework, with the goal of creating an open library of \glspl{NCS} to share with the community.
At the moment, the interactions of \gls{CLEAVE} and tools such as Docker are still quite superficial.
Our goal is to achieve a much tighter integration, e.g.\ by providing the toolkit as pre-packaged container images.
Finally, the validity of the results obtained by the framework will have to be verified through more thorough, realistic scenarios than what we have been able to show in this work.
In particular, we intend to perform large-scale experimentation targetting 5G cellular deployments, as this technology is set to become the backbone of edge networks in the near future.

### 2022-12-19 publications/2022CLEAVE/experiments.tex
glossary fixes

In this section, we demonstrate the utility of \gls{CLEAVE} by walking readers through an example use case.
With this, we aim to showcase the ability of our framework to provide accurate and repeatable measurements of the performance of \acp{NCS} deployed on edge computing infrastructure.

@@

We believe this to be an interesting and representative use-case scenario for a number of reasons.
Firstly, the inverted pendulum is broadly used as a benchmark in control-system literature, see for instance\ \cite{Baumann2018LowPower} and\ \cite{Natale2004InvPendEthernet}.
A key reason for this is that such a relatively simplistic system allows for straightforward reparametrization to obtain varied system dynamics, which in turn makes a broader range of experiments possible.
Secondly, similar, albeit non-Edge-enabled, setups have been a reality in industry for almost two decades, as evidenced by\ \cite{Gupta2010NCSOverview}.
Thirdly, although real deployments would most assuredly employ mobile technologies due to the added benefits of mobility and range, the \acp{RTT} offered by these technologies are higher, or at best equal, those offered by WiFi.
4G, for instance, has average \acp{RTT} of around a few hundred milliseconds, which is orders of magnitude greater than the values measured in our scenario, and current 5G deployments offer \gls{RTT} values barely matching the sub-\SI{10}{\milli\second} values we obtain for WiFi.
Thus, we argue that using WiFi allows for more experimental freedom, as the system can be tweaked to obtain worse \acp{RTT} more akin to those of mobile technologies while still having the option to study best-case scenarios with very low-latencies.
Finally, video analytics is one of the main proposed use cases for edge computing~\cite{Ananthanarayanan2017Analytics,Yi2017Analytics,Wang2018Analytics}, and thus we foresee edge \gls{NCS} deployments being deployed in parallel with such applications in the future.
However, before deployment can become a reality, key questions such as: 
``what are the baseline requirements of the \gls{NCS}?''; ``what are the achievable best-case end-to-end latencies in the system?''; and
``how does the concurrent video-streaming traffic affect \gls{NCS} stability?'' need to be answered.

@@

\begin{figure}[t]
    \centering
    \begin{subfigure}[t]{\columnwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_single_loop_toppled}
        \caption{Fraction of toppled plants.}\label{fig:single:topple}
    \end{subfigure}\\
    \begin{subfigure}[t]{\columnwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_single_loop_rms}
        \caption{Mean angle \acs*{RMS}; red line indicates \( y = 180 \).}\label{fig:single:rms}
    \end{subfigure}\\
    \begin{subfigure}[t]{\columnwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_single_loop_rtts}
        \caption{
            Mean \acsp*{RTT}.
        }\label{fig:single:rtt}
    \end{subfigure}%
    \caption[caption]{
        Baseline results.
        Error bars indicate \SI{95}{\percent} \acp{CI}.
        }%
    \label{fig:single}
\end{figure}

@@

As expected, higher sampling rates tend to correlate with better quality of control; at higher sampling rates the system was able to reach stability at higher \acp{RTT}.
These initial results already hint at interesting consequences for such an edge-bound \gls{NCS} deployment.
For instance, it is clear from \cref{fig:single:topple,fig:single:rms} that network delays can, to a certain extent, be compensated for by increasing the sampling rate of the system.
A corollary of this is, conversely, that at lower network latencies \acp{NCS} are able to stabilize at lower sampling rates.
Adaptive sampling might thus be a viable method for optimizing resource utilization.

Once baselines for single loops in the use case scenario have been established, we can proceed to studying the interaction between the \acp{NCS} and the video-streaming applications.
We deploy \num{6} control loops on the experimental setup depicted in \cref{fig:cleave:expsetup}.
On the remaining \num{4} clients we run the \emph{iperf3}\footnote{iperf3: \url{https://iperf.fr/}} traffic load generator, each generating \SI[per-mode=symbol]{6.5}{\mega\bit\per\second} of uplink \gls{UDP} traffic.
This emulates the load generated on the network by \SI{1080}{p} Full-HD video streaming, originating from the clients and terminating in the cloudlet.
Based on the baseline results, we execute setups with \gls{NCS} plant sampling rates of \SIlist{20;40;60}{\hertz}.
Each sampling rate configuration is run for \SI{5}{\minute}, and then repeated \num{30} times to obtain statistical significance.
Once again, repetitions of this scenario are executed automatically in batches using a simple Python script and \emph{docker-py}.

\begin{figure}[t]
    \centering
    \begin{subfigure}[h]{.22\textwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_video_topple_frac}
        \caption{Toppled plants.}\label{fig:video:toppled}
    \end{subfigure}%
    \hfill%
    \begin{subfigure}[h]{.22\textwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_video_angle_rms}
        \caption{Angle \gls{RMS}.}\label{fig:video:rms}
    \end{subfigure}\\
    \begin{subfigure}[h]{.22\textwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_video_drop_frac}
        \caption{Packet losses.}\label{fig:video:drop}
    \end{subfigure}%
    \hfill%
    \begin{subfigure}[h]{.22\textwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_video_rtt}
        \caption{\acsp{RTT}.}\label{fig:video:rtt}
    \end{subfigure}%
    \caption{
        Multi-loop, resource-constrained setup results.
        Error bars indicate \SI{95}{\percent} \acp{CI} in all plots.
    }\label{fig:video:results}
\end{figure}

@@

\cref{fig:video:drop,fig:video:rtt} explains these behaviors.
Whereas both the \SIlist{20;40}{\hertz} setups show losses well below \SI{5}{\percent}, the \SI{60}{\hertz} scenario shows an average of around \SI{13}{\percent} of datagrams lost.
The differences in \acp{RTT} results are equally telling; \acp{RTT} for the \SI{60}{\hertz} setup were on average approx.\ \num{3} times those for the \SIlist{20;40}{\hertz} setups.
These results stem from the contention for network resources, and hint at important trade-offs system designers will have to take into consideration when designing and developing \acp{NCS} for deployment on the Edge.
The Edge will be \emph{multi-tenant} and \emph{multi-instance}. 
\gls{NCS} deployments will have to be designed with shared resources in mind, and given the complexity of these systems, experimental tools like \gls{CLEAVE} will be key for their succesful adoption and massification.

>>>>

In this section, we demonstrate the utility of \gls{CLEAVE} by walking readers through an example use case.
With this, we aim to showcase the ability of our framework to provide accurate and repeatable measurements of the performance of \glspl{NCS} deployed on edge computing infrastructure.

@@

We believe this to be an interesting and representative use-case scenario for a number of reasons.
Firstly, the inverted pendulum is broadly used as a benchmark in control-system literature, see for instance\ \cite{Baumann2018LowPower} and\ \cite{Natale2004InvPendEthernet}.
A key reason for this is that such a relatively simplistic system allows for straightforward reparametrization to obtain varied system dynamics, which in turn makes a broader range of experiments possible.
Secondly, similar, albeit non-Edge-enabled, setups have been a reality in industry for almost two decades, as evidenced by\ \cite{Gupta2010NCSOverview}.
Thirdly, although real deployments would most assuredly employ mobile technologies due to the added benefits of mobility and range, the \glspl{RTT} offered by these technologies are higher, or at best equal, those offered by WiFi.
4G, for instance, has average \glspl{RTT} of around a few hundred milliseconds, which is orders of magnitude greater than the values measured in our scenario, and current 5G deployments offer \gls{RTT} values barely matching the sub-\SI{10}{\milli\second} values we obtain for WiFi.
Thus, we argue that using WiFi allows for more experimental freedom, as the system can be tweaked to obtain worse \glspl{RTT} more akin to those of mobile technologies while still having the option to study best-case scenarios with very low-latencies.
Finally, video analytics is one of the main proposed use cases for edge computing~\cite{Ananthanarayanan2017Analytics,Yi2017Analytics,Wang2018Analytics}, and thus we foresee edge \gls{NCS} deployments being deployed in parallel with such applications in the future.
However, before deployment can become a reality, key questions such as: 
``what are the baseline requirements of the \gls{NCS}?''; ``what are the achievable best-case end-to-end latencies in the system?''; and
``how does the concurrent video-streaming traffic affect \gls{NCS} stability?'' need to be answered.

@@

\begin{figure}[t]
    \centering
    \begin{subfigure}[t]{\columnwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_single_loop_toppled}
        \caption{Fraction of toppled plants.}\label{fig:single:topple}
    \end{subfigure}\\
    \begin{subfigure}[t]{\columnwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_single_loop_rms}
        \caption{Mean angle \gls{RMS}; red line indicates \( y = 180 \).}\label{fig:single:rms}
    \end{subfigure}\\
    \begin{subfigure}[t]{\columnwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_single_loop_rtts}
        \caption{
            Mean \glspl{RTT}.
        }\label{fig:single:rtt}
    \end{subfigure}%
    \caption[caption]{
        Baseline results.
        Error bars indicate \SI{95}{\percent} \glspl{CI}.
        }%
    \label{fig:single}
\end{figure}

@@

As expected, higher sampling rates tend to correlate with better quality of control; at higher sampling rates the system was able to reach stability at higher \glspl{RTT}.
These initial results already hint at interesting consequences for such an edge-bound \gls{NCS} deployment.
For instance, it is clear from \cref{fig:single:topple,fig:single:rms} that network delays can, to a certain extent, be compensated for by increasing the sampling rate of the system.
A corollary of this is, conversely, that at lower network latencies \glspl{NCS} are able to stabilize at lower sampling rates.
Adaptive sampling might thus be a viable method for optimizing resource utilization.

Once baselines for single loops in the use case scenario have been established, we can proceed to studying the interaction between the \glspl{NCS} and the video-streaming applications.
We deploy \num{6} control loops on the experimental setup depicted in \cref{fig:cleave:expsetup}.
On the remaining \num{4} clients we run the \emph{iperf3}\footnote{iperf3: \url{https://iperf.fr/}} traffic load generator, each generating \SI[per-mode=symbol]{6.5}{\mega\bit\per\second} of uplink \gls{UDP} traffic.
This emulates the load generated on the network by \SI{1080}{p} Full-HD video streaming, originating from the clients and terminating in the cloudlet.
Based on the baseline results, we execute setups with \gls{NCS} plant sampling rates of \SIlist{20;40;60}{\hertz}.
Each sampling rate configuration is run for \SI{5}{\minute}, and then repeated \num{30} times to obtain statistical significance.
Once again, repetitions of this scenario are executed automatically in batches using a simple Python script and \emph{docker-py}.

\begin{figure}[t]
    \centering
    \begin{subfigure}[h]{.22\textwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_video_topple_frac}
        \caption{Toppled plants.}\label{fig:video:toppled}
    \end{subfigure}%
    \hfill%
    \begin{subfigure}[h]{.22\textwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_video_angle_rms}
        \caption{Angle \gls{RMS}.}\label{fig:video:rms}
    \end{subfigure}\\
    \begin{subfigure}[h]{.22\textwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_video_drop_frac}
        \caption{Packet losses.}\label{fig:video:drop}
    \end{subfigure}%
    \hfill%
    \begin{subfigure}[h]{.22\textwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_video_rtt}
        \caption{\glspl{RTT}.}\label{fig:video:rtt}
    \end{subfigure}%
    \caption{
        Multi-loop, resource-constrained setup results.
        Error bars indicate \SI{95}{\percent} \glspl{CI} in all plots.
    }\label{fig:video:results}
\end{figure}

@@

\cref{fig:video:drop,fig:video:rtt} explains these behaviors.
Whereas both the \SIlist{20;40}{\hertz} setups show losses well below \SI{5}{\percent}, the \SI{60}{\hertz} scenario shows an average of around \SI{13}{\percent} of datagrams lost.
The differences in \glspl{RTT} results are equally telling; \glspl{RTT} for the \SI{60}{\hertz} setup were on average approx.\ \num{3} times those for the \SIlist{20;40}{\hertz} setups.
These results stem from the contention for network resources, and hint at important trade-offs system designers will have to take into consideration when designing and developing \glspl{NCS} for deployment on the Edge.
The Edge will be \emph{multi-tenant} and \emph{multi-instance}. 
\gls{NCS} deployments will have to be designed with shared resources in mind, and given the complexity of these systems, experimental tools like \gls{CLEAVE} will be key for their succesful adoption and massification.
